---
title: "GLMER analysis pipeline — Hemispheric Dissociation via Attentional Isolation and tRNS"
author: "Michele Tosi"
output: html_document
---

# Overview

Generalised linear mixed-effects models (binomial family) on trial-level
Multiple Object Tracking (MOT) responses (correct = 1, incorrect = 0),
with by-subject random intercepts and, where appropriate, by-subject
random slopes for `VisualF`. All post-hoc contrasts are Holm-corrected
via `emmeans`.

Set `DERIV` below to the folder produced by the `preprocessing/` scripts.

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(readxl)
library(lme4)
library(car)
library(emmeans)
library(rstatix)
library(DHARMa)
library(performance)
library(lmtest)
```

```{r config}
# Point this at the derivatives folder produced by preprocessing/
DERIV <- "../derivatives"
```

# 1. Load data

```{r load}
my_data       <- read.csv(file.path(DERIV, "long_format_collapsed.csv"))
dt            <- read.csv(file.path(DERIV, "long_format_single_trial.csv"))
dt_iso        <- read.csv(file.path(DERIV, "all_isolation_time_split_single_trial.csv"))
my_data_delta <- read.csv(file.path(DERIV, "delta_long_format_averaged.csv"))
```

# 2. Outlier removal at baseline

Outliers are detected on Pre-session accuracy within each `Collection`
(montage cohort) × `VisualF` cell using the standard 1.5×IQR rule
(`rstatix::identify_outliers`). Flagged subjects are removed from every
downstream dataset.

```{r outliers}
outdect <- my_data %>%
  filter(Session == "Pre") %>%
  group_by(Collection, VisualF) %>%
  identify_outliers(Response)

drop_ids <- unique(outdect$Subject)

dt            <- dt            %>% filter(!(Subject %in% drop_ids))
my_data       <- my_data       %>% filter(!(Subject %in% drop_ids))
my_data_delta <- my_data_delta %>% filter(!(Subject %in% drop_ids))
dt_iso        <- dt_iso        %>% filter(!(Subject %in% drop_ids))

# Focus on Pre / Post_1 (Post_2 = 10-min retest, secondary)
dt_filt <- dt %>% filter(Session != "Post_2")
dt_filt$manipulation <- NULL

my_data       <- my_data       %>% filter(Session != "Post_2")
my_data_delta <- my_data_delta %>% filter(Session != "P2")
my_data_delta <- my_data_delta %>% filter(Manipulation != "Control")
my_data_delta$manipulation <- NULL
```

# 3. Baseline check — no pre-test difference between groups

Confirms randomisation worked: pre-test accuracy should not differ by
Manipulation or Stimulation before the isolation window.

```{r baseline-check}
fit_0 <- glmer(
  Response ~ Manipulation * Stimulation + (1 + VisualF | Subject),
  family  = binomial,
  data    = dt_filt,
  subset  = (Session == "Pre" & Collection == 2),
  control = glmerControl(optimizer = "bobyqa")
)
summary(fit_0)
Anova(fit_0, type = 3)
```

# 4. Model selection

Fit a family of candidate random-effects structures, then compare with
likelihood-ratio tests (`anova`). The winning model has a by-subject
random intercept + random slope for `VisualF`.

```{r model-selection, eval=FALSE}
fit_alt1 <- glmer(Response ~ Session * Manipulation * Stimulation +
                    (1 + VisualF | Subject),
                  family = binomial, data = dt_filt,
                  control = glmerControl(optimizer = "bobyqa"))

fit_alt2 <- glmer(Response ~ Session * Manipulation * Stimulation * VisualF +
                    (1 | Subject),
                  family = binomial, data = dt_filt,
                  control = glmerControl(optimizer = "bobyqa"))

fit_alt3 <- glmer(Response ~ Session * Manipulation * Stimulation +
                    (1 | Subject),
                  family = binomial, data = dt_filt,
                  control = glmerControl(optimizer = "bobyqa"))

fit_alt4 <- glmer(Response ~ Session * Manipulation * Stimulation +
                    (1 + Stimulation | Subject),
                  family = binomial, data = dt_filt,
                  control = glmerControl(optimizer = "bobyqa"))

anova(fit_alt1, fit_alt2, fit_alt3, fit_alt4)
```

# 5. Main effect: attentional gain in the ignored visual field

Test the three-way `Session × Manipulation × Stimulation` interaction
(Experiment 1 = right montage, Collection == 1).

```{r main-model}
fit_1 <- glmer(
  Response ~ Session * Manipulation * Stimulation + (1 + VisualF | Subject),
  family  = binomial,
  data    = dt_filt,
  subset  = (Collection == "1"),
  control = glmerControl(optimizer = "bobyqa")
)
summary(fit_1)
Anova(fit_1, type = 3)

emm <- emmeans(fit_1, ~ Session : Manipulation : Stimulation)
pairs(emm, adjust = "Holm")
```

# 6. Control check — no tRNS effect in the absence of isolation

Bilateral tracking group, contralateral (left) visual field, sham vs
active. A null result here means tRNS alone (without manipulation) is
not driving the effect.

```{r control-check}
fit_control <- glmer(
  Response ~ Session * Stimulation * VisualF + (1 | Subject),
  family  = binomial,
  data    = dt_filt,
  subset  = (Collection == "1" & Manipulation == "Control"),
  control = glmerControl(optimizer = "bobyqa")
)
Anova(fit_control, type = 3)
```

# 7. Pre / During / Post trajectory in the ATTENDED visual field

Combine Pre / Post_1 blocks with the During-isolation trials, then
model accuracy across the three time points to show that the
manipulation phase produces a within-session boost that returns to
baseline at Post.

```{r pre-during-post}
dt_iso["Session"] <- "During"
dt_pre_post <- dt %>% filter(Manipulation != "Ignored Visual Field")

# Harmonise column names between the two sources
colnames(dt_iso)[which(colnames(dt_iso) == "Time.Split")] <- "Time_Split"
colnames(dt_iso)[which(colnames(dt_iso) == "Isolation")]  <- "Manipulation"
dt_iso$Manipulation[dt_iso$Manipulation == "Bilateral_Control"] <- "Control"
dt_iso$Collection <- as.numeric(as.character(dt_iso$Collection))
dt_iso$Time_Split <- as.character(dt_iso$Time_Split)

dt_iso_full <- dt_iso %>% select(-Time_Split)

custom_order <- c("Subject", "Manipulation", "Stimulation", "Session",
                  "VisualF", "Response", "Collection")
dt_iso_full <- dt_iso_full[, custom_order]
dt_pre_post <- dt_pre_post[, intersect(custom_order, colnames(dt_pre_post))]

dt_pre_during_post <- dplyr::bind_rows(dt_pre_post, dt_iso_full)

fit_3 <- glmer(
  Response ~ Session * Stimulation * Manipulation + (1 + VisualF | Subject),
  family  = binomial,
  data    = dt_pre_during_post,
  subset  = (Collection == "1" & Session != "Post_2"),
  control = glmerControl(optimizer = "bobyqa")
)
Anova(fit_3, type = 3)
pairs(emmeans(fit_3, ~ Session), adjust = "Holm")
```

# 8. Within-window dynamics — 10-min bins

Split the 30-minute isolation into three 10-minute bins to test whether
accuracy drifts across the window (e.g. fatigue effects).

```{r time-bins}
fit_4 <- glmer(
  Response ~ Manipulation * Stimulation * Time_Split + (1 + VisualF | Subject),
  family  = binomial,
  data    = dt_iso,
  subset  = (Collection == "2" & Session != "Post_2"),
  control = glmerControl(optimizer = "bobyqa")
)
Anova(fit_4, type = 3)
pairs(emmeans(fit_4, ~ Time_Split), adjust = "Holm")
```

# 9. Left vs right montage — Δ-accuracy in the ignored visual field

Compare the pre-to-post gain (`delta_accuracy`) between the two
experiments in the ignored visual field only. This is the key
hemispheric-dissociation test (linear mixed model on subject-level Δ).

```{r delta-lmm}
my_data_delta_ign <- my_data_delta %>%
  filter(Manipulation == "Ignored Visual Field", Session == "P1")

fit_7 <- lmer(delta_accuracy ~ Collection * Stimulation + (1 | Subject),
              data = my_data_delta_ign)
summary(fit_7)
Anova(fit_7, type = 3)

emm <- emmeans(fit_7, ~ Collection : Stimulation)
pairs(emm, adjust = "holm")
eff_size(emm, sigma = sigma(fit_7), edf = df.residual(fit_7))
```

# 10. Ipsilateral vs contralateral to the stimulation site

Recode `VisualF` into `Ipsilateral` / `Contralateral` relative to
stimulation, then fit a three-way `Stimulation × Collection ×
visual_stim` GLMER on the attended-side trials during isolation.

```{r ipsi-contra}
dt_iso_full <- dt_iso_full %>%
  mutate(visual_stim = case_when(
    Collection == 1 & VisualF == "Right" ~ "Ipsilateral",
    Collection == 1 & VisualF == "Left"  ~ "Contralateral",
    Collection == 2 & VisualF == "Right" ~ "Contralateral",
    Collection == 2 & VisualF == "Left"  ~ "Ipsilateral",
    TRUE ~ "Other"
  ))

fit_8 <- glmer(
  Response ~ Stimulation * Collection * visual_stim + (1 | Subject),
  family  = binomial,
  data    = dt_iso_full,
  subset  = Manipulation == "Attended Visual Field",
  control = glmerControl(optimizer = "bobyqa")
)
Anova(fit_8, type = 3)

emm <- emmeans(fit_8, ~ Stimulation : Collection : visual_stim)
pairs(emm, adjust = "Holm")
```
