---
title: "Plots — Hemispheric Dissociation via Attentional Isolation and tRNS"
author: "Michele Tosi"
output: html_document
---

Reproduces the paper figures (Figure 2, 3, 4, 5) from the cleaned
derivatives. All bars are means with `mean_se` error bars, individual
subjects overlaid.

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(lme4)
library(car)
library(emmeans)
library(rstatix)
library(gridExtra)
library(ggsignif)
```

```{r config}
DERIV   <- "../derivatives"
OUT_DIR <- "../figures/generated"
dir.create(OUT_DIR, recursive = TRUE, showWarnings = FALSE)
```

# Load data + drop baseline outliers (same rule as analysis script)

```{r load}
my_data       <- read.csv(file.path(DERIV, "long_format_collapsed.csv"))
dt            <- read.csv(file.path(DERIV, "long_format_single_trial.csv"))
dt_iso        <- read.csv(file.path(DERIV, "all_isolation_time_split_single_trial.csv"))
my_data_delta <- read.csv(file.path(DERIV, "delta_long_format_averaged.csv"))

outdect <- my_data %>%
  filter(Session == "Pre") %>%
  group_by(Collection, VisualF) %>%
  identify_outliers(Response)
drop_ids <- unique(outdect$Subject)

dt            <- dt            %>% filter(!(Subject %in% drop_ids))
my_data       <- my_data       %>% filter(!(Subject %in% drop_ids))
my_data_delta <- my_data_delta %>% filter(!(Subject %in% drop_ids))
dt_iso        <- dt_iso        %>% filter(!(Subject %in% drop_ids))

dt_filt <- dt %>% filter(Session != "Post_2")
my_data <- my_data %>% filter(Session != "Post_2")
my_data_delta <- my_data_delta %>% filter(Session != "P2")
```

# Reusable theme

```{r theme}
theme_paper <- function(base = 18) {
  theme(
    strip.background = element_blank(),
    panel.background = element_blank(),
    panel.grid       = element_blank(),
    axis.line        = element_line(color = "black", size = 1.5),
    axis.ticks       = element_line(color = "black"),
    axis.text        = element_text(size = base, face = "bold", color = "black"),
    axis.title       = element_text(size = base, face = "bold", vjust = 2),
    axis.title.y     = element_text(size = base, face = "bold",
                                    margin = margin(r = 12)),
    plot.title       = element_text(size = base, hjust = 0.5),
    legend.text      = element_text(size = base + 2),
    legend.title     = element_text(size = base + 2),
    legend.key.size  = unit(0.6, "cm"),
    strip.text.x     = element_text(size = base + 4, hjust = 0.5)
  )
}
```

# Merge Pre / During / Post into a single frame (for Fig. 2 & 3)

```{r merge-pdp}
dt_iso["Session"] <- "During"
dt_pre_post <- dt %>% filter(Manipulation != "Ignored Visual Field")

colnames(dt_iso)[which(colnames(dt_iso) == "Time.Split")] <- "Time_Split"
colnames(dt_iso)[which(colnames(dt_iso) == "Isolation")]  <- "Manipulation"
dt_iso$Manipulation[dt_iso$Manipulation == "Bilateral_Control"] <- "Control"
dt_iso$Collection <- as.numeric(as.character(dt_iso$Collection))
dt_iso$Time_Split <- as.character(dt_iso$Time_Split)
dt_iso_full <- dt_iso %>% select(-Time_Split)

keep_cols <- c("Subject", "Manipulation", "Stimulation", "Session",
               "VisualF", "Response", "Collection")
dt_iso_full <- dt_iso_full[, keep_cols]
dt_pre_post <- dt_pre_post[, intersect(keep_cols, colnames(dt_pre_post))]

dt_pre_during_post <- dplyr::bind_rows(dt_pre_post, dt_iso_full)
dt_pre_during_post$Session[dt_pre_during_post$Session == "Post_1"] <- "Post"
dt_pre_during_post$Session <- factor(dt_pre_during_post$Session,
                                     levels = c("Pre", "During", "Post"))

my_data_pdp <- dt_pre_during_post %>%
  group_by(Subject, Manipulation, Stimulation, Session, Collection, VisualF) %>%
  summarise(Response = mean(Response, na.rm = TRUE), .groups = "drop")

# Cosmetic recoding for the main pre/post plot
my_data$Manipulation[my_data$Manipulation == "Ignored Visual Field"]  <- "Ignored"
my_data$Manipulation[my_data$Manipulation == "Attended Visual Field"] <- "Attended"
my_data$Manipulation <- factor(my_data$Manipulation,
                               levels = c("Ignored", "Attended", "Control"))
my_data$Session[my_data$Session == "Post_1"] <- "Post"
my_data$Session <- factor(my_data$Session, levels = c("Pre", "Post"))
my_data$Stimulation <- factor(my_data$Stimulation, levels = c("Sham", "Active"))

my_data_delta$Manipulation[my_data_delta$Manipulation == "Ignored Visual Field"]  <- "Ignored"
my_data_delta$Manipulation[my_data_delta$Manipulation == "Attended Visual Field"] <- "Attended"
my_data_delta$Manipulation <- factor(my_data_delta$Manipulation,
                                     levels = c("Ignored", "Attended", "Control"))
my_data_delta$Stimulation <- factor(my_data_delta$Stimulation,
                                    levels = c("Sham", "Active"))
```

# Figure 2 — Pre vs Post accuracy, right montage (Experiment 1)

```{r fig2, fig.height=10, fig.width=14}
p_fig2 <- ggplot(my_data %>% filter(Collection == "1"),
                 aes(x = Session, fill = Manipulation, y = Response * 100)) +
  geom_bar(position = position_dodge(), stat = "summary", fun = mean,
           width = 0.7, colour = "black") +
  scale_fill_manual(values = c("#0072B2", "#E69F00", "#009E73")) +
  facet_wrap(Stimulation ~ Manipulation, scales = "free_x") +
  stat_summary(geom = "errorbar", fun.data = mean_se,
               position = position_dodge(), width = 0.3, size = 1) +
  geom_point(shape = 21, size = 2.5, color = "black") +
  labs(x = "Session", y = "MOT Accuracy (%)") +
  coord_cartesian(ylim = c(40, 95)) +
  scale_y_continuous(breaks = seq(40, 95, by = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.85) +
  theme_paper()

ggsave(file.path(OUT_DIR, "Figure2_right_montage.tiff"),
       plot = p_fig2, width = 17.6, height = 17.6, units = "cm",
       dpi = 300, device = "tiff", compression = "lzw")
p_fig2
```

# Figure 3 — Pre vs Post accuracy, left montage (Experiment 2)

```{r fig3, fig.height=10, fig.width=14}
p_fig3 <- ggplot(my_data %>% filter(Collection == "2" | Manipulation == "Control"),
                 aes(x = Session, fill = Manipulation, y = Response * 100)) +
  geom_bar(position = position_dodge(), stat = "summary", fun = mean,
           width = 0.7, colour = "black") +
  scale_fill_manual(values = c("#0072B2", "#E69F00", "#009E73")) +
  facet_wrap(Stimulation ~ Manipulation, scales = "free_x") +
  stat_summary(geom = "errorbar", fun.data = mean_se,
               position = position_dodge(), width = 0.3, size = 1) +
  geom_point(shape = 21, size = 2.5, color = "black") +
  labs(x = "Session", y = "MOT Accuracy (%)") +
  coord_cartesian(ylim = c(40, 95)) +
  scale_y_continuous(breaks = seq(40, 95, by = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.85) +
  theme_paper()

ggsave(file.path(OUT_DIR, "Figure3_left_montage.tiff"),
       plot = p_fig3, width = 17.6, height = 17.6, units = "cm",
       dpi = 300, device = "tiff", compression = "lzw")
p_fig3
```

# Figure 4 — Δ (Post − Pre) MOT accuracy in the ignored visual field

Compares the pre-to-post gain in the ignored visual field between sham
and active tRNS, separately per montage. This is the plot that carries
the paper's headline hemispheric-dissociation result.

```{r fig4, fig.height=7, fig.width=7}
delta_ign <- my_data_delta %>%
  filter(Manipulation == "Ignored", Session == "P1")
delta_ign$Collection <- factor(delta_ign$Collection, levels = c("2", "1"))
custom_labels <- c("1" = "Right Montage", "2" = "Left Montage")

p_fig4 <- ggplot(delta_ign, aes(x = Stimulation, fill = Stimulation,
                                y = delta_accuracy * 100)) +
  geom_bar(position = position_dodge(), stat = "summary", fun = mean,
           width = 0.55, colour = "black") +
  scale_fill_manual(values = c("#1B9E77", "#E6AB02")) +
  facet_wrap(~Collection, scales = "free_x",
             labeller = labeller(Collection = custom_labels)) +
  stat_summary(geom = "errorbar", fun.data = mean_se,
               position = position_dodge(), width = 0.3) +
  geom_point(alpha = 0.7, shape = 21, size = 3, color = "black") +
  labs(x = "Stimulation", y = "Change MOT Accuracy (%)") +
  coord_cartesian(ylim = c(-30, 30)) +
  scale_y_continuous(breaks = seq(-30, 30, by = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.85) +
  theme_paper()

ggsave(file.path(OUT_DIR, "Figure4_delta_ignored.tiff"),
       plot = p_fig4, width = 17.6, height = 17.6, units = "cm",
       dpi = 300, device = "tiff", compression = "lzw")
p_fig4
```

# Figure 5 — Ipsi vs contralateral (during-isolation, attended side)

Compute the active-minus-sham delta separately for the ipsilateral and
contralateral hemifield relative to the stimulated hemisphere, during
the 30-min isolation window (attended side only). This is what makes
the double dissociation between left and right montages visible.

```{r fig5, fig.height=8, fig.width=12}
dt_iso_full <- dt_iso_full %>%
  mutate(visual_stim = case_when(
    Collection == 1 & VisualF == "Right" ~ "Ipsilateral",
    Collection == 1 & VisualF == "Left"  ~ "Contralateral",
    Collection == 2 & VisualF == "Right" ~ "Contralateral",
    Collection == 2 & VisualF == "Left"  ~ "Ipsilateral",
    TRUE ~ "Other"
  ))

dt_iso_collapsed <- dt_iso_full %>%
  group_by(Subject, Manipulation, Stimulation, Session, Collection,
           VisualF, visual_stim) %>%
  summarise(Response = mean(Response, na.rm = TRUE), .groups = "drop")

delta_difference <- dt_iso_collapsed %>%
  group_by(Subject, Session, Manipulation, Collection, visual_stim, VisualF) %>%
  filter(n() == 2) %>%
  summarise(Delta = Response[Stimulation == "Active"] -
                    Response[Stimulation == "Sham"],
            .groups = "drop")

delta_difference$Collection <- factor(delta_difference$Collection,
                                      levels = c("2", "1"))

p_fig5 <- ggplot(delta_difference %>%
                   filter(Manipulation == "Attended Visual Field"),
                 aes(x = visual_stim, fill = visual_stim, y = Delta * 100)) +
  geom_bar(position = position_dodge(width = 0.85), stat = "summary",
           fun = mean, width = 0.4, colour = "black") +
  scale_fill_manual(values = c("#D3D3D3", "#333333")) +
  facet_wrap(~Collection, scales = "free_x",
             labeller = labeller(Collection = custom_labels)) +
  stat_summary(geom = "errorbar", fun.data = mean_se,
               position = position_dodge(width = 0.85), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = 0.85),
             shape = 21, size = 3, color = "black") +
  labs(x = "Isolation side", y = "% Change MOT accuracy") +
  coord_cartesian(ylim = c(-25, 20)) +
  scale_y_continuous(breaks = seq(-25, 20, by = 5)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.85) +
  theme_paper()

ggsave(file.path(OUT_DIR, "Figure5_hemisphere_dissociation.tiff"),
       plot = p_fig5, width = 17.6, height = 12, units = "cm",
       dpi = 300, device = "tiff", compression = "lzw")
p_fig5
```
